HealpixNet: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2018-08-15 scnn.tex
\cdot => \times

The lowest possible resolution is given by the base partitioning of the surface into $N_{pix} = 12$ equal-sized pixels (right-most sphere in \figref{pooling}).
As each pixel is subdivided in four, the second coarser resolution is $N_{pix} = N_{side}^2 \cdot 12 = 2^2 \cdot 12 = 48$ pixels (middle sphere in \figref{pooling}), the third is $N_{pix} = N_{side}^2 \cdot 12 = 4^2 \cdot 12 = 192$ pixels, etc., where $N_{side} = 1, 2, 4, 8, \ldots$ is the grid resolution parameter.
High-resolutions HEALPix maps easily reach millions of pixels.
\figref{example_maps} shows two examples of HEALPix maps: the Cosmic Microwave Background \citep{planck2015overview}, galaxies found in Sloan Digital Sky Survey Data Release 14 \citep{abolfathi2017sdssDR14}, and an example simulated mass map on the footprint of Dark Energy Survey Data Release 1 \citep{des2018dr1}.

@@

From the HEALPix pixelization, we build a weighted undirected graph $\G = (\V, \E, \W)$, where $\V$ is the set of $N_{pix} = |\V|$ nodes, $\E$ is the set of edges, and $\W$ is the weighted adjacency matrix.
In our graph, each pixel $i$ is represented by a node (also called vertex) $v_i \in \V$.
Each node $v_i$ is then connected to the $8$ (or $7$)\footnote{\label{neighbors}The $12 \cdot 4 = 48$ pixels at the corner of each rhombus of the base dodecahedron only have 7 neighboring pixels. See Figures~\ref{fig:healpix_graph_4} and~\ref{fig:pooling}.} nodes $v_j$ which represent the neighboring pixels $j$ of pixel $i$, forming edges $(v_i, v_j) \in \E$. Given those edges, we define the weighted adjacency matrix $\W \in \R^{N_{pix} \times N_{pix}}$ as
\begin{equation*}
	\W_{ij} = \begin{cases}
		\exp \left( -\frac{\|\x_i-\x_j\|_2^2}{\rho^2} \right) & \text{if pixels $i$ and $j$ are neighbors,} \\
		0 & \text{otherwise,} \\
	\end{cases}
\end{equation*}
where $\x_i$ is a vector encoding the 3-dimensional coordinates of pixel $i$, and
\begin{equation*}
	\rho = \frac{1}{|\E|} \sum_{(v_i, v_j) \in \E} \|\x_i-\x_j\|_2
\end{equation*}
is the average Euclidean distance over all connected pixels. This weighting scheme is important as distances between pixels are not equal.
Other weighting schemes are possible. For example,~\cite{khasanova2017graphomni} uses the inverse of the distance instead. We found out that the one proposed above works well for our purpose, and did not investigate other approaches, leaving it to future work.
\figref{healpix_graph_4} shows a graph constructed from the HEALPix sampling of a sphere.

@@

Given a matrix $\X = [\x_1, \ldots, \x_{F_{in}}] \in \R^{N_{pix} \times F_{in}}$, a GC layer computes $\Y = GC(\X) = [\y_1, \ldots, \y_{F_{out}}] \in \R^{N_{pix} \times F_{out}}$, where $N_{pix}$ is the number of pixels (and nodes), $F_{in}$ is the number of input features, and $F_{out}$ is the number of output features.
Using the efficient graph convolution from \eqnref{graph_convolution_cheby}, each output feature map is computed as
\begin{equation*}
	\y_i = \sum_{j=1}^{F_{in}} h_{\theta_{ij}}(\tL) \x_j + b_i \in \R^{N_{pix}}, \ \ \forall i \in [F_{out}].
\end{equation*}
As such, a GC layer is composed of $F_{in} \cdot F_{out}$ filters, each parameterized by $K$ numbers (see \secref{efficient_convolution}). A bias term $\b b \in \R^{F_{out}}$ is jointly optimized.
% only if batch normalization is not used

@@

A FC layer takes as input $\x \in \R^{F_{in}}$ and outputs
\begin{equation*}
	\y = FC(\x) = \b W \x + \b b \in \R^{F_{out}} ,
\end{equation*}
where $\W \in \R^{F_{out} \times F_{in}}$ and $\b b \in \R^{F_{out}}$ are the parameters to be learned.
Note that the output $\Y \in \R^{N_{pix} \times F_{out}}$ of the last GC (or the output $\Y \in \R^{N_{stat} \times F_{out}}$ of the ST), is vectorized as $\x = \vect(\X) \in \R^{F_{in}}$ before being fed to the first FC, where $F_{in} = N_{pix} \cdot F_{out}$ (or $N_{stat} \cdot F_{out}$).

>>>>

The lowest possible resolution is given by the base partitioning of the surface into $N_{pix} = 12$ equal-sized pixels (right-most sphere in \figref{pooling}).
As each pixel is subdivided in four, the second coarser resolution is $N_{pix} = N_{side}^2 \times 12 = 2^2 \times 12 = 48$ pixels (middle sphere in \figref{pooling}), the third is $N_{pix} = N_{side}^2 \times 12 = 4^2 \times 12 = 192$ pixels, etc., where $N_{side} = 1, 2, 4, 8, \ldots$ is the grid resolution parameter.
High-resolutions HEALPix maps easily reach millions of pixels.
\figref{example_maps} shows two examples of HEALPix maps: the Cosmic Microwave Background \citep{planck2015overview}, galaxies found in Sloan Digital Sky Survey Data Release 14 \citep{abolfathi2017sdssDR14}, and an example simulated mass map on the footprint of Dark Energy Survey Data Release 1 \citep{des2018dr1}.

@@

From the HEALPix pixelization, we build a weighted undirected graph $\G = (\V, \E, \W)$, where $\V$ is the set of $N_{pix} = |\V|$ nodes, $\E$ is the set of edges, and $\W$ is the weighted adjacency matrix.
In our graph, each pixel $i$ is represented by a node (also called vertex) $v_i \in \V$.
Each node $v_i$ is then connected to the $8$ (or $7$)\footnote{\label{neighbors}The $12 \times 4 = 48$ pixels at the corner of each rhombus of the base dodecahedron only have 7 neighboring pixels. See Figures~\ref{fig:healpix_graph_4} and~\ref{fig:pooling}.} nodes $v_j$ which represent the neighboring pixels $j$ of pixel $i$, forming edges $(v_i, v_j) \in \E$. Given those edges, we define the weighted adjacency matrix $\W \in \R^{N_{pix} \times N_{pix}}$ as
\begin{equation*}
	\W_{ij} = \begin{cases}
		\exp \left( -\frac{\|\x_i-\x_j\|_2^2}{\rho^2} \right) & \text{if pixels $i$ and $j$ are neighbors,} \\
		0 & \text{otherwise,} \\
	\end{cases}
\end{equation*}
where $\x_i$ is a vector encoding the 3-dimensional coordinates of pixel $i$, and
\begin{equation*}
	\rho = \frac{1}{|\E|} \sum_{(v_i, v_j) \in \E} \|\x_i-\x_j\|_2
\end{equation*}
is the average Euclidean distance over all connected pixels. This weighting scheme is important as distances between pixels are not equal.
Other weighting schemes are possible. For example,~\cite{khasanova2017graphomni} uses the inverse of the distance instead. We found out that the one proposed above works well for our purpose, and did not investigate other approaches, leaving it to future work.
\figref{healpix_graph_4} shows a graph constructed from the HEALPix sampling of a sphere.

@@

Given a matrix $\X = [\x_1, \ldots, \x_{F_{in}}] \in \R^{N_{pix} \times F_{in}}$, a GC layer computes $\Y = GC(\X) = [\y_1, \ldots, \y_{F_{out}}] \in \R^{N_{pix} \times F_{out}}$, where $N_{pix}$ is the number of pixels (and nodes), $F_{in}$ is the number of input features, and $F_{out}$ is the number of output features.
Using the efficient graph convolution from \eqnref{graph_convolution_cheby}, each output feature map is computed as
\begin{equation*}
	\y_i = \sum_{j=1}^{F_{in}} h_{\theta_{ij}}(\tL) \x_j + b_i \in \R^{N_{pix}}, \ \ \forall i \in [F_{out}].
\end{equation*}
As such, a GC layer is composed of $F_{in} \times F_{out}$ filters, each parameterized by $K$ numbers (see \secref{efficient_convolution}). A bias term $\b b \in \R^{F_{out}}$ is jointly optimized.
% only if batch normalization is not used

@@

A FC layer takes as input $\x \in \R^{F_{in}}$ and outputs
\begin{equation*}
	\y = FC(\x) = \b W \x + \b b \in \R^{F_{out}} ,
\end{equation*}
where $\W \in \R^{F_{out} \times F_{in}}$ and $\b b \in \R^{F_{out}}$ are the parameters to be learned.
Note that the output $\Y \in \R^{N_{pix} \times F_{out}}$ of the last GC (or the output $\Y \in \R^{N_{stat} \times F_{out}}$ of the ST), is vectorized as $\x = \vect(\X) \in \R^{F_{in}}$ before being fed to the first FC, where $F_{in} = N_{pix} \times F_{out}$ (or $N_{stat} \times F_{out}$).
